---
title: "fact_table"
author: "Mokeli Pelamoko"
date: "April 10, 2016"
output: html_document
---
## 1. Loading the dataset

```{r}
library(dplyr)
library(ggplot2)
library(data.table)  
library(tidyr)
library(reshape2)
options( java.parameters = "-Xmx4g" )
""
reservations <- fread("3865_reservations.csv",check.names = T,
                      header = T)
reservations$loyalty_user_id <- factor(reservations$loyalty_user_id)
reservations$loyalty_reservation_id <-factor(reservations$loyalty_reservation_id)
reservations$made_on <- as.POSIXct(reservations$made_on,origin="1970-01-01",tz="GMT")
reservations$reservation_date <-as.POSIXct(reservations$reservation_date,
                                           origin="1970-01-01",tz="GMT")

pos <- fread("3865_pos.csv",check.names = T,header=T)
pos$created_on_utc <- as.POSIXct(pos$created_on_utc, origin="1970-01-01",tz="GMT")
pos$modified_on_utc <- as.POSIXct(pos$modified_on_utc, origin="1970-01-01",tz="GMT")
pos$open_date <- as.POSIXct(pos$open_date,origin = "1970-01-01", tz = "GMT")
pos$loyalty_visit_id <- as.factor(pos$loyalty_visit_id)
pos$loyalty_reservation_id <- as.factor(pos$loyalty_reservation_id)
pos$loyalty_user_id <- as.factor(pos$loyalty_user_id)
pos$server_id <- as.factor(pos$server_id)

pos_items <- fread("pos_items_final.csv",check.names = T, header = T)
pos_items$amount <- as.numeric(pos_items$amount)
pos_items$amount_per_unit <- as.numeric(pos_items$amount_per_unit)
pos_items$discount <- as.numeric(pos_items$discount)
pos_items$quantity <- as.numeric(pos_items$quantity)

pos_items$loyalty_visit_id <- as.factor(pos_items$loyalty_visit_id)
pos_items$loyalty_reservation_id <- as.factor(pos_items$loyalty_reservation_id)
pos_items$loyalty_visit_item_id <- as.factor(pos_items$loyalty_visit_item_id)

pos_items$item_category_name <- as.factor(pos_items$item_category_name)
pos_items$item_name <- as.factor(pos_items$item_name)
pos_items$item_subcategory_name <- as.factor(pos_items$item_subcategory_name)
```

## Test. Testing w/ Unique Users 

This section was just used to test and determine data integrity problems

```{r}
x <- master_fact %>% 
      filter(loyalty_user_id %in% c("3216887","18821897","16696")) %>%
      arrange(open_date)
y <- x %>% arrange(open_date) %>%
     select(loyalty_user_id,loyalty_reservation_id,open_date) %>%
     group_by(loyalty_user_id) %>%
     mutate(diffx = difftime(open_date,lag(open_date,default=NA),units="days")) %>%
     group_by(loyalty_user_id) %>%
     dplyr::summarise(ya= mean(as.double(diffx),na.rm = T))
```

## 2. Merging Reservations with POS

I merged it first, because there's  a lot of junk in POS. Thus, only keeping POS lines that match with a reservation from Open Table. Removing the any reservation with an 0 id and the two outliers. Also, there are some lines that duplicates even after merging, so I am removing any lines with same reservation id, user id, discount amount, visit id, and spend amount. 

```{r}
rsvp_pos <- merge(reservations,pos,by="loyalty_reservation_id")
rsvp_pos$loyalty_user_id.y <- NULL
names(rsvp_pos)[4] <- "loyalty_user_id"
duplicates_flag <-duplicated(rsvp_pos,by=c("loyalty_reservation_id","loyalty_user_id",
                                           "discount","loyalty_visit_id","spend"))
rsvp_pos <- rsvp_pos[!duplicates_flag,]
rsvp_pos <- rsvp_pos %>% 
                filter(!loyalty_user_id %in% c('12484585','14324743','0') & 
                        loyalty_reservation_id != '0')
```

## 3. Creating Fact Table from joined RSVP & POS

Creating the lowest grain fact table. The visit id is not one-to-one with reservation id like Dan told us. There are some transactions where there are two visit id with different spend and / or discount but with the same open date and reservation number. Thus, I am summing spend and discount by loyalty user id and reservation id. The rest of the lines just create various dimensions like day of the week,etc
```{r}
master_fact <- rsvp_pos %>%
            group_by(loyalty_user_id,loyalty_reservation_id) %>%
            summarise(party_size=min(party_size),
                      rsvp_server_name=min(server_name.x),
                      rsvp_table=min(table_name.x),                      
                      pos_server_name=min(server_name.y),
                      pos_table=min(table_name.y),
                      state=min(state),
                      cover_count=max(cover_count),
                      discount=sum(discount),   
                      spend=sum(spend),
                      made_on=min(made_on),
                      rsvp_date=min(reservation_date),
                      open_date=min(open_date))

master_fact$elapsed_days_madeon_open <- round(with(master_fact,
                                           difftime(open_date,made_on,
                                                     units = "days")),4)
master_fact$elapsed_mins_rsvp_open <-round(with(master_fact,
                                          difftime(open_date,rsvp_date,
                                                   units="mins")),2)

master_fact$day_visit <- factor( weekdays(master_fact$open_date),
                             levels = c("Sunday","Monday","Tuesday", "Wednesday",
                                        "Thursday","Friday","Saturday"))

master_fact$year_visit <- year(master_fact$open_date)

master_fact$month_visit <- month(master_fact$open_date)

master_fact$day_of_month_visit <- as.integer(format(master_fact$open_date,"%d"))

master_fact$weekday_visit <- with(master_fact,ifelse(day_visit %in% c("Monday","Tuesday",
                                                                "Wednesday","Thursday"), 1,0))
master_fact$weekend_visit <- with(master_fact,ifelse(day_visit %in% c("Friday","Saturday",
                                                                "Sunday"), 1,0))

master_fact$hour_visit <- as.numeric(format(master_fact$open_date,"%H"))

master_fact$breakfast_visit <-with(master_fact,ifelse(!(day_visit %in% c("Saturday","Sunday")) & 
                                                    between(hour_visit,8,11),1,0))
master_fact$lunch_visit <-with(master_fact,ifelse(!(day_visit %in% c("Saturday","Sunday")) & 
                                                    between(hour_visit,11,16),1,0))

master_fact$dinner_visit <-with(master_fact,ifelse(between(hour_visit,17,24),1,0))

master_fact$brunch_visit <-with(master_fact,ifelse(day_visit %in% c("Saturday","Sunday") & 
                                                     between(hour_visit,11,14),1,0))

```

### 4. Adding holiday flags to Master Fact Table
Christmas is any day ? 5 from December 25th
New Year's is December 31st or January 1st
Valentine's is any day ? 5 from February 14th
Mother's Day is the Second Saturday or Sunday in May
```{r}

#takes years in the dataset as input and returns Mother's Day for each of the year
fun_mothers_day <- function(years){
  may_first <- as.Date(paste0(years,"-05-01"))
  mothers_day <- c()
  for(i in 1:length(may_first)){
    days <- seq(from=may_first[i],by="day",length.out = 14)
    sundays <- days[weekdays(days)=="Sunday"]
    mothers_day[i] <- as.character(max(sundays))
  }
  return(mothers_day)
}

master_fact$christmas_flag <-with(master_fact,
                                  ifelse(month_visit==12 & between(day_of_month_visit,20,30),1,0))

master_fact$newyears_flag <- with(master_fact,ifelse((month_visit == 12 & 
                                                day_of_month_visit == 31) |
                                                (month_visit == 1 & day_of_month_visit == 1),1,0))
master_fact$valentine_flag <-with(master_fact,
                                  ifelse(month_visit==2 & between(day_of_month_visit,9,19),1,0))

years <- levels(as.factor(master_fact$year_visit))
all_mothers_days <- as.Date(fun_mothers_day(years))

master_fact$mother_day <- 0
for(i in 1:length(all_mothers_days))
  master_fact$mother_day <- with(master_fact,ifelse(year_visit ==year(all_mothers_days[i]),
                                 mday(all_mothers_days[i]),mother_day))
master_fact$mother_day_flag <- with(master_fact,ifelse(month_visit==5 & 
                                                         between(day_of_month_visit,mother_day-1,
                                                               mother_day),1,0))

```


### 5. Creating aggregates by user from RSVP & POS joined tbl

```{r}
fact_aggregates <- master_fact %>% 
             group_by(loyalty_user_id) %>%
             summarise(num_of_repeats = n(),
                       min_party_size = min(party_size), 
                       max_party_size = max(party_size),
                       avg_party_size = mean(party_size),
                       min_cover_count = min(cover_count), 
                       max_cover_count = max(cover_count),
                       avg_cover_count = mean(cover_count),
                       min_discount = min(discount), 
                       max_discount = max(discount),
                       avg_discount = mean(discount),                      
                       min_spend = min(spend), 
                       max_spend = max(spend),
                       avg_spend = mean(spend), 
                       min_diffdays_madeon_open = min(elapsed_days_madeon_open), 
                       max_diffdays_madeon_open = max(elapsed_days_madeon_open),
                       avg_diffdays_madeon_open = mean(elapsed_days_madeon_open),
                       min_diffmins_rsvp_open = min(elapsed_mins_rsvp_open), 
                       max_diffmins_rsvp_open = max(elapsed_mins_rsvp_open),
                       avg_diffmins_rsvp_open = mean(elapsed_mins_rsvp_open),
                       avg_weekday_visits = mean(weekday_visit),
                       avg_weekend_visits = mean(weekend_visit),
                       avg_breakfast_visits= mean(breakfast_visit),
                       avg_lunch_visits = mean(lunch_visit),
                       avg_dinner_visits = mean(dinner_visit),
                       avg_brunch_visits = mean(brunch_visit),
                       avg_christmas_visits = mean(christmas_flag),
                       avg_newyears_visits = mean(newyears_flag),
                       avg_valentine_visits = mean(valentine_flag),
                       avg_mothers_day_visits = mean(mother_day_flag)
                       )
fact_aggregates$has.repeated <- ifelse(fact_aggregates$num_of_repeats > 1,"yes","no")
```

### 6. Creating dimensions to obtain first and last visit from RSVP & POS items

```{r}
fact_first <- master_fact %>% 
             group_by(loyalty_user_id) %>%
             filter(open_date == min(open_date)) %>%
             summarise(first_party_size = party_size,
                       first_server_name = pos_server_name,
                       first_table = pos_table,
                       first_discount = discount,
                       first_spend = spend,
                       first_visit_date = open_date,
                       first_diffdays_madeon_visit = elapsed_days_madeon_open,
                       first_diffmins_rsvp_open = elapsed_mins_rsvp_open,
                       first_weekday_visit = weekday_visit,
                       first_weekend_visit = weekend_visit,
                       first_breakfast = breakfast_visit,
                       first_lunch_visit = lunch_visit,
                       first_dinner_visit = dinner_visit,
                       first_brunch_visit = brunch_visit,
                       first_christmas_visit = christmas_flag,
                       first_newyears_visit = newyears_flag,
                       first_valentine_visit = valentine_flag,
                       first_mother_day_flag = mother_day_flag)

fact_last <- master_fact %>% 
             group_by(loyalty_user_id) %>%
             filter(open_date == max(open_date)) %>%
             summarise(last_party_size = party_size,
                       last_server_name = pos_server_name,
                       last_table = pos_table,
                       last_discount = discount,
                       last_spend = spend,
                       last_visit_date = open_date,
                       last_diffdays_madeon_visit = elapsed_days_madeon_open,
                       last_diffmins_rsvp_open = elapsed_mins_rsvp_open,
                       last_weekday_visit = weekday_visit,
                       last_weekend_visit = weekend_visit,
                       last_breakfast_visit = breakfast_visit,
                       last_lunch_visit = lunch_visit,
                       last_dinner_visit = dinner_visit,
                       last_brunch_visit = brunch_visit,
                       last_christmas_visit = christmas_flag,
                       last_newyears_visit = newyears_flag,
                       last_valentine_visit = valentine_flag,
                       last_mother_day_flag = mother_day_flag)
```

### 7. Creating dimensions about the server from RSVP & POS items
```{r}
fact_server_source <- master_fact %>% 
                       group_by(loyalty_user_id,pos_server_name) %>%
                       summarise(n = n()) %>%
                       mutate(max_same_server = max(n))

fact_server_repeats <- fact_server_source %>%
                    select(loyalty_user_id, max_same_server) %>%
                    distinct(loyalty_user_id,
                             max_same_server)

fact_server_counts <- fact_server_source %>%
                      group_by(loyalty_user_id) %>%
                      summarise(num_of_different_servers=n())
fact_server <- merge(fact_server_counts,fact_server_repeats,by="loyalty_user_id")

rm(fact_server_source,fact_server_repeats,fact_server_counts)
```


### 8. Creating dimensions about the table from RSVP & POS items
```{r}
fact_table_source <- master_fact %>% 
                       group_by(loyalty_user_id,pos_table) %>%
                       summarise(n = n()) %>%
                       mutate(max_same_table = max(n))

fact_table_repeats <- fact_table_source %>%
                    select(loyalty_user_id,max_same_table) %>%
                    distinct(loyalty_user_id,max_same_table)

fact_table_counts <- fact_table_source %>%
                      group_by(loyalty_user_id) %>%
                      summarise(num_of_different_tables=n())
fact_table <- merge(fact_table_counts,fact_table_repeats,by="loyalty_user_id")
rm(fact_table_source,fact_table_repeats,fact_table_counts)
```

## 9. Combining aggregates, first, last, server, and table

```{r}

rsvp_pos_main <- Reduce(function(...) merge(...,by="loyalty_user_id"),
                    list(fact_aggregates,fact_first,fact_last,fact_server,fact_table))

rm(fact_aggregates,fact_first,fact_last,fact_server,fact_table)

```


## 10. Creating POS Items Dimensions

Creating average loyalty category by loyalty user id. 
```{r}

items_rsvp <- pos_items %>% 
                 filter(loyalty_reservation_id != "0" & 
                         !(loyalty_category_name %in% c("","loyalty_category_name"))) %>%
                 group_by(loyalty_reservation_id,
                          loyalty_category_name) %>%
                 summarise(amount = sum(amount)) %>%
                 spread(loyalty_category_name,amount)
setnames(items_rsvp, make.names(colnames(items_rsvp)))
items_rsvp[is.na(items_rsvp)] <- 0

items_user <- merge(select(master_fact,loyalty_reservation_id,loyalty_user_id),
                    items_rsvp,by="loyalty_reservation_id") 

items_aggregates <- items_user %>%
                    group_by(loyalty_user_id) %>%
                    summarise_each(funs(mean(.,na.rm=TRUE)),3:ncol(items_user))
names(items_aggregates)[-1] <- paste0("avg_",names(items_aggregates)[-1],"_spend")

rm(items_rsvp,items_user)

```

### 11. Creating dimensions from Code and State
How do we treat blanks? 
I used rsvp_pos for code to ensure that the cancelled and no shows are counted. Otherwise, there are more reservations than in pos. 
```{r}
rsvp_code <- rsvp_pos %>%
                group_by(loyalty_user_id,loyalty_reservation_id) %>%
                summarise(code = min(code)) %>%
                mutate(code=ifelse(code %in% c("Anniversary","Birthday"),
                                   code,"Other"),
                       value = 1) %>%
                group_by(loyalty_user_id,code) %>%
                summarise(value = sum(value)) %>%
                spread(code,value)
rsvp_code[is.na(rsvp_code)] <- 0
names(rsvp_code)[-1] <- paste0(names(rsvp_code)[-1],"_visits")

rsvp_state <- reservations %>%
              filter(ordinal==1 & state %in% c("Cancelled","No Show")) %>%
              dplyr::select(loyalty_user_id,loyalty_reservation_id,state) %>%
              mutate(value=1) %>%
              group_by(loyalty_user_id,state) %>%
              summarise(value=sum(value)) %>%
              spread(state,value)
setnames(rsvp_state, make.names(colnames(rsvp_state)))
names(rsvp_state)[-1] <- paste0(names(rsvp_state)[-1],"_visits")

rm(rsvp_pos)

```

### 12. Creating Repeat Visits and Average Days b/t Visis dimensions
Average days between visit is lag difference between reservation day by loyalty user id after the open date as been sorted.

The repeat visit on the same day is created by combining the month and day of each visit, converting it to an integer, then taking the difference. Any loyalty user with 0 difference would indicate that the user repeated on the same date.

```{r}

bt_visits <- master_fact %>% 
             arrange(open_date) %>%
             select(loyalty_user_id,loyalty_reservation_id,open_date,
                    month_visit,day_of_month_visit) %>%
             mutate(cmb_month_day = as.integer(paste0(month_visit,day_of_month_visit))) %>%
             group_by(loyalty_user_id) %>%
             mutate(days_bt_visits = difftime(open_date,lag(open_date,default=NA),
                                     units="days")) 

avg_bt_visit <- bt_visits %>%
                 group_by(loyalty_user_id) %>%
                 dplyr::summarise(avg_days_bt_visits = 
                                    round(mean(as.double(days_bt_visits),
                                         na.rm = T),3))
avg_bt_visit[is.na(avg_bt_visit)] <- 0
flag_same_day <-duplicated(bt_visits,by=c("loyalty_user_id","cmb_month_day"))
                                   
same_day_visit <- bt_visits[flag_same_day,]   %>% 
                  group_by(loyalty_user_id) %>%
                  summarise(same_day_visits =n())
                  
repeat_time_bt_visits <- merge(avg_bt_visit,same_day_visit,
                               by="loyalty_user_id",all.x = T,all.y = F)
repeat_time_bt_visits[is.na(repeat_time_bt_visits)] <- 0
rm(bt_visits,avg_bt_visit,flag_same_day,same_day_visit)
```

## 12. Combining POS Items, Code, State, and Repeat/Average days tables with reservation and pos talbe

```{r}

fact_table <- merge(rsvp_pos_main,items_aggregates,by="loyalty_user_id")
fact_table <- merge(fact_table,rsvp_code,by="loyalty_user_id")
fact_table <- merge(fact_table,rsvp_state,all.x = TRUE,
                    all.y=FALSE,by="loyalty_user_id")
fact_table <- merge(fact_table,repeat_time_bt_visits,all.x = TRUE,
                    all.y=FALSE,by="loyalty_user_id")
fact_table[is.na(fact_table)] <- 0

```


