---
title: "Practicum"
author: "Mokeli P, Nisha I, & Rachel"
date: "February 14, 2016"
output: html_document
---

# Loading the Dataset and converting to proper time format
```{r}
library(dplyr)
library(ggplot2)
reservations <- read.csv("3865_reservations.csv")[,c(1,3,4,5,7,8,9,10,11,12,13,14)]
reservations$loyalty_user_id <- factor(reservations$loyalty_user_id)
reservations$loyalty_reservation_id <-factor(reservations$loyalty_reservation_id)
reservations$made_on <- as.POSIXct(reservations$made_on,origin="1970-01-01",z="GMT")
reservations$reservation_date <-as.POSIXct(reservations$reservation_date,
                                           origin="1970-01-01",z="GMT")

pos <- read.csv("3865_pos.csv")[,c(2,4:8,10:17)]
pos_items <- read.csv("pos_items.csv")

```

# Removing all ordinals except ordinal equal 1

```{r}
str(reservations)
table(reservations$ordinal)
#Keeping only ordinal equaling 1
reservations <- subset(reservations,reservations$ordinal ==1)
```

We may have to revisit and keep the highest ordinal in order to retain code information.

# Reservation Codes
```{r}
summary.codes <- reservations %>% 
                 select(code) %>% 
                 group_by(code) %>% 
                 summarise(n=n()) %>%
                 arrange(-n)
```
Summary of Reservations

# Table of Loyalty User
```{r}
user.visits <- reservations %>% 
               select(loyalty_user_id) %>%
               group_by(loyalty_user_id) %>%
               summarise(visits =n()) %>%
               arrange(-visits)
```
Users 12484585 and 14324743 have over 1400 reservations. Is this in error?

# Parent Venue
```{r}
table(reservations$parent_venue_id)
```
There's only 1 parent idea

# Creating Bins for Party Size
```{r}
summary(reservations$party_size)
ggplot(data=reservations,aes(x=party_size,fill="blue")) + geom_histogram(binwidth = 1)
table(reservations$party_size)

sizebreaks <- c(18,25,35,45,55,65,10000000000)
sizelabels = c("18-24","25-34","35-44","45-54","55-64","65+")
test_set <- test_set %>%
                 mutate(age_bucket= cut(age, breaks= agebreaks, right= FALSE, labels= agelabels))

```

```{r}

```

